---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# chmloader

<!-- badges: start -->
<!-- badges: end -->

The goal of chmloader is to download the Canopy Height Model (CHM) data from 
this recent work by [Tolan et al. (2023)](https://research.facebook.com/blog/2023/4/every-tree-counts-large-scale-mapping-of-canopy-height-at-the-resolution-of-individual-trees/)
The data is downloaded from aws s3 storage - further details on the bucket can
be found [here](https://registry.opendata.aws/dataforgood-fb-forests/)

## Installation

You can install the development version of chmloader like so:

``` r
pak::pkg_install("chmloader")
```

## Example

This is a basic example which shows you how to download some data:

```{r example}
library(chmloader)

sundarbans <- sf::st_point(c(89.2, 22.0)) |>
  sf::st_sfc(crs = 4326) |>
  sf::st_buffer(5000)

st <- Sys.time()
sundarbans_chm <- download_chm(sundarbans)
round(difftime(Sys.time(), st, units = "secs"))

sdb_chm <- terra::rast(sundarbans_chm)
sdb_chm[sdb_chm < 1] <- NA
terra::plot(sdb_chm,
  col = hcl.colors(256, "viridis")
)
```

This package has only the one function `download_chm` which downloads the CHM
data from the aws s3 storage. The function takes a `sf` object as input and
returns a character vector for the output raster file path. The download uses
gdalwarp (via `sf::gdal_utils`) to efficiently retrieve only the required data 
from across multiple tiles - the default resolution is 1 m but this can be 
reprojected on the fly as needed. 